Safe Transferable Regions

### 2016-10-22 paper/type-system.tex
Type safety stmt edited. Transfer safety removed.

Replacing $\pi$'s and $\rho$'s with $\loc$'s at runtime results in a
new class of expressions that are not captured in
Fig.~\ref{fig:fb-syntax}. Appendix contains full syntax that includes
such expressions. To help state the type safety theorem, we define the
syntactic class of (runtime) values:
\begin{smathpar}
\begin{array}{lclcl}
v & \in & \mathtt{values} & \coloneqq & \C{new}\;
    B\inang{\tbar}\inang{\overline{\loc}}(\bar{v}) \ALT
\lambdaexp{\loc}{\rhobar\,|\,\phi}{\taubar \; \xbar}{e} \ALT
\C{new}\; \RgnZT{\toploc\loc}(v)
%  &     & & & \C{new}\; \RgnZT{\rgn}(v)\\
\end{array}
\end{smathpar}
The first two forms are obtained by using locations for region
annotations in $\C{new}$ and lambda expressions. The last form is the
value that $\C{new}\;\RgnZ$ expressions gets evaluated to; the
location $\toploc$ stands for the region $\toprgn$, and $\loc$ is the
location of the newly allocated transferable region.  The following
type safety theorem shows that a well-typed program will never attempt
to dereference an ``invalid'' reference (a reference to an object in a
region that has been transferred or freed):
\begin{theorem}
\emph{(\textbf{Type Safety})}
\label{thm:fb-type-safety}
$\forall e, \tau, \rhoenv, \rhomap, \rgn$, such that $\rgn \in
\rhoenv$ and $\tywf{\rhoenv}{\phicx}$, if $\hastyp{\emptyA,\rgn,
\cdot}{e}{\tau}$, then either $e$ is a value, or $e$ raises an
exception ($\redstoo{(e,\rhomap)}{\invalidexn}$), or there exists an
$e'$ and a $\rhomap'$ such that $\redstoo{(e,\rhomap)}{(e',\rhomap')}$
and $\hastyp{\emptyASigp,\rgn, \cdot}{e'}{\tau}$.
\end{theorem}

Furthermore, we prove the following theorem about \FB, which, in
conjunction with the type safety theorem, implies the safety of region
transfers across address spaces:
\begin{theorem}
\emph{(\textbf{Transfer Safety})}
\label{thm:fb-transfer-safety}
$\forall$ $v$, $\rhoenv$, $\rhoenv'$, $\rhomap$, $\rhomap'$, $\phicx$,
$\phicx'$, $\rgn$, $\rgn'$, such that $\rgn \in \rhoenv$, $\rgn' \in
\rhoenv'$, $\rgn_i \neq \toprgn$ and
$\rgn_i \notin dom(\rhomap') \cup \rhoenv'$, if $\hastyp{\emptyA,\rgn,
\cdot}{\C{new}\; \RgnZT{\rgn_i}(v)}{\RgnZT{\toprgn}}$, then
$\hastyp{(\rhomap'[\rgn_i \mapsto \CLOSED],\rhoenv',\cdot,\phicx')}
{\\\C{new}\; \RgnZT{\rgn_i}(v)}{\RgnZT{\toprgn}}$
\end{theorem}
The above theorem states that if a $\C{new}\; \RgnZT{\rgn_i}(v)$ value
is well-typed under one context, then it is also well-typed under
every other context, whose $\rhomap$ maps $\rgn_i$ to closed
($\CLOSED$) state.  Thus, a recipient of a transferable region only
needs to add a binding for the region to its $\rhomap$ in order to
preserve its type safety. 
% Notably, this result could not have been
% established if it were possible for a transferable region to contain
% references to objects outside the region.

>>>>

Replacing $\pi$'s and $\rho$'s with $\loc$'s at runtime results in a
new class of expressions that are not captured in
Fig.~\ref{fig:fb-syntax}. Appendix contains full syntax that includes
such expressions. To help state the type safety theorem, we define the
syntactic class of (runtime) values:
\begin{smathpar}
\begin{array}{lclcl}
v & \in & \mathtt{values} & \coloneqq & \C{new}\;
    B\inang{\tbar}\inang{\overline{\loc}}(\bar{v}) \ALT
\lambdaexp{\loc}{\rhobar\,|\,\phi}{\taubar \; \xbar}{e} \ALT
\C{new}\; \RgnZT{\toploc\loc}(v)
%  &     & & & \C{new}\; \RgnZT{\rgn}(v)\\
\end{array}
\end{smathpar}
The first two forms are obtained by using locations for region
annotations in $\C{new}$ and lambda expressions. The last form is the
value that $\C{new}\;\RgnZ$ expressions gets evaluated to; the
location $\toploc$ stands for the region $\toprgn$, and $\loc$ is the
location of the newly allocated transferable region.  The following
type safety theorem shows that a well-typed program will never attempt
to dereference an ``invalid'' reference (a reference to an object in a
region that has been transferred or freed):
\begin{theorem}
\emph{(\textbf{Type Safety})}
\label{thm:fb-type-safety}
$\forall e, \tau, \Delta, \mem$, such that $\tywf{\rhoenv}{\phicx}$, if $\hastyp{\emptyA,
\cdot}{e}{\tau}$, then either $e$ is a value, or $e$ raises an
exception ($\redstoo{\Delta}{(e,\mem)}{\invalidexn}$), or there exists an
$e'$ and a $\mem'$ such that $\redstoo{\Delta}{(e,\mem)}{(e',\mem')}$
and $\hastyp{\emptyASigp,\cdot}{e'}{\tau}$.
\end{theorem}

% Furthermore, we prove the following theorem about \FB, which, in
% conjunction with the type safety theorem, implies the safety of region
% transfers across address spaces:
% \begin{theorem}
% \emph{(\textbf{Transfer Safety})}
% \label{thm:fb-transfer-safety}
% $\forall$ $v$, $\rhoenv$, $\rhoenv'$, $\mem$, $\mem'$, $\phicx$,
% $\phicx'$, such that $\loc \not\in dom(\mem')$, if $\hastyp{\emptyA, \cdot}{\C{new}\;
% \RgnZT{\toploc\loc}(v)}{\RgnZT{\toprgn}}$, then
% $\hastyp{\rhoenv',\cdot,\phicx')}
% {\\\C{new}\; \RgnZT{\rgn_i}(v)}{\RgnZT{\toprgn}}$
% \end{theorem}
% The above theorem states that if a $\C{new}\; \RgnZT{\rgn_i}(v)$ value
% is well-typed under one context, then it is also well-typed under
% every other context, whose $\mem$ maps $\rgn_i$ to closed
% ($\CLOSED$) state.  Thus, a recipient of a transferable region only
% needs to add a binding for the region to its $\mem$ in order to
% preserve its type safety. 
% Notably, this result could not have been
% established if it were possible for a transferable region to contain
% references to objects outside the region.
